---
title: "Reproducible Research - Assignment 1"
author: "Michal Buczynski"
date: "Saturday, January 17, 2015"
output:
  html_document:
    keep_md: yes
---
### Loading and preprocessing the data

```{r, echo=TRUE}


setwd("D:/Data Scientist Specialisation/Reproducible research/repdata_data_activity")

activity_data <- read.csv("activity.csv", header=TRUE)

str(activity_data)


```

### What is mean total number of steps taken per day?
```{r, echo=TRUE}

hist_data <- tapply(activity_data$steps, activity_data$date, sum)

hist(hist_data, col="blue", xlab="Number of steps", 
     main="Histogram")
grid()


myMean <- tapply(activity_data$steps, activity_data$date, function(x) mean(x,na.rm=TRUE))
myMedian <- tapply(activity_data$steps, activity_data$date, function(x) median(x, na.rm=TRUE))

myMedian
round(myMean,2)
```

### What is the average daily activity pattern?
```{r, echo=TRUE}

avgDaily <- tapply(activity_data$steps, activity_data$interval, 
                   function(x) mean(x,na.rm=TRUE))


plot(unique(activity_data$interval), avgDaily, col="blue", type="l",
     xlab="5 - minute intervals", ylab="Avg number of steps taken, across all days",
     xlim=c(0,2355))

max <- which.max(avgDaily)


```

`r max` time interval has the average maximum number of steps across all days in the dataset.

### Imputing missing values
Calculate  of NAs in the dataset
```{r, echo=TRUE}

any(is.na(activity_data))
all(is.na(activity_data))

sumNA <- sum(is.na(activity_data$steps))
```
### There are `r sumNA` missing values in the dataset.

Data imputation - replace NA value with average 5 minute interval steps taken across all days.create data with interval and correspodning average steps taken across all days.

```{r, echo=TRUE}
nrow(activity_data)
myData <- data.frame(unique(activity_data$interval), as.numeric(avgDaily))
names(myData) <- c("interval", "steps_avg")

for (i in 1:nrow(activity_data)) {
  
    if(is.na(activity_data[i,1])) {
      
          for (j in 1:288) {
            
                if(activity_data[i,3] == myData[j,1]) {
                  
                  activity_data[i,1] <- myData[j,2]
                  
                }
          } 
    } 
}

new_activity_data <- activity_data

new_hist_data <- tapply(new_activity_data$steps, new_activity_data$date, sum)


hist(new_hist_data, col="blue", xlab="Number of steps", 
     main="Histogram")


# Mean and median

new_myMean <- tapply(new_activity_data$steps, new_activity_data$date, mean)
new_myMedian <- tapply(new_activity_data$steps, new_activity_data$date, median)

round(new_myMean,2)
round(new_myMedian,2)
```
The dataset differs significantly after the imputation of NA values. 

### Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE, fig.width=15}

new_activity_data$date <- as.Date(new_activity_data$date)

wDay <- c()

for (i in 1:17568) {
  
  if(weekdays(new_activity_data[i,2]) == "sobota" | 
       weekdays(new_activity_data[i,2]) == "niedziela") {

        wDay[i] <- 1 # Weekend
      
      } else {  
        
        wDay[i] <-  2 # Working day
            
      }
}

new_activity_data <- data.frame(new_activity_data,wDay)

new_activity_data$wDay <- factor(new_activity_data$wDay, levels = c(1,2),
                                 labels=c("Weekend", "Working day"))

split_data <- split(new_activity_data, new_activity_data$wDay)

avgDailyWeekend <- tapply(split_data[[1]]$steps, split_data[[1]]$interval, 
                mean)

avgDailyWorking <- tapply(split_data[[2]]$steps, split_data[[2]]$interval, 
                          mean)

par(mfrow = c(1,2))

plot(unique(split_data[[1]]$interval), avgDailyWeekend, type="l", col="red",
     main="Weekend", xlab="5 - minute intervals", 
     ylab="Avg number of steps taken across all weekend days")
plot(unique(split_data[[2]]$interval), avgDailyWorking, type="l", col="blue",
     main="Working day",  xlab="5 - minute intervals", 
     ylab="Avg number of steps taken across all working days")


```

knit2html(input = "PA1_template.Rmd")
